---
title: Bibliometric analysis of the term "Anthropocene"
author: Daniel Knitter, Katja Augustin, Ehsan Biniyaz, Wolfgang Hamer, Michael Kuhwald, Malte Schwanebeck, Rainer Duttmann
date: "`r Sys.Date()`"
bibliography: references.bib
output: md_document
---

Data was downloaded on Di 19. Dez 11:51:54 CET 2017 from [Web of Science](http://apps.webofknowledge.com/summary.do?product=WOS&search_mode=GeneralSearch&qid=1&SID=F2uyu8Tz33gg81ezRRV&page=) using query "anthropocene"

# bibliometric analyses

Data preparation and reshaping is conducted using the `bibliometrix` package [@aria2017]. 

```{r bibliometrix}
library(magrittr)
## devtools::install_github("massimoaria/bibliometrix")
library(bibliometrix)

raw_data <- readFiles("./data/wos_query_anthropocene/wos_anthropocene_topic_1.bib",
                      "./data/wos_query_anthropocene/wos_anthropocene_topic_2.bib",
                      "./data/wos_query_anthropocene/wos_anthropocene_topic_3.bib",
                      "./data/wos_query_anthropocene/wos_anthropocene_topic_4.bib")
raw_df <- convert2df(file = raw_data, dbsource = "isi", format = "bibtex")

results <- biblioAnalysis(raw_df, sep = ";")

summary(object = results, k = 10, pause = FALSE)
plot(x = results, k = 10, pause = FALSE)

# Create keyword co-occurrencies network
NetMatrix <- biblioNetwork(M = raw_df, analysis = "co-occurrences", network = "keywords", sep = ";")
# Plot the network
net <- networkPlot(NetMatrix, n = 20, Title = "Keyword Co-occurrences", type = "kamada", size=T)
```

```{r tibble}
library(magrittr)
library(dplyr)
library(readr)
wok_field_tags <- read_tsv("data/wos_tags.txt", skip = 1)
wok_field_tags %>%
  filter(Tag == "SC")

library(tidyr)
raw_tbl <- as.tbl(raw_df)

ResearchArea <- raw_tbl %>%  
  mutate(SC = strsplit(as.character(SC), ";")) %>%  
  unnest %>%
  mutate(SC = trimws(x = SC, which = "left")) %>%  
  mutate(SC = as.factor(SC)) %>%
  group_by(SC) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

library(ggplot2)
library(ggrepel)

ra_plot <- ResearchArea %>%
  na.omit %>%  
  mutate(SC = factor(SC, levels = SC[rev(order(count))])) %>%  
  ggplot(mapping = aes(x = SC,y = count)) +
  geom_point() +
  theme_minimal() +  
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + 
  geom_label_repel(
    data = subset(x = ResearchArea, subset = ResearchArea$count > 50),
    aes(label = SC),
    size = 5,
    box.padding = 0.25,
    point.padding = 0.3
  ) + 
  labs(title = "Research Areas resulting from \"Anthropocene\" topic search",
       subtitle = "data from Web of Science; Topic Article search term fields are: Title, Abstract, Keywords")
ra_plot

ggsave(filename = "ResearchArea_plot.png",
       plot = ra_plot,
       scale = 2,
       width = 15,
       height = 10,
       dpi = 300,
       device = "png",
       units = "cm")
```

The analysis is conducted using R [@R] and the contributed packages `rmarkdown` [@rmarkdown], `magrittr` [@magrittr], `bibliometrix` [@aria2017], `dplyr` [@dplyr], `readr` [@readr], `tidyr`  [@tidyr], `ggplot2` [@ggplot2] and  `ggrepel` [@ggrepel].

# References

<div id="refs"></div>

# Colophon

```{r }
devtools::session_info()
```
